Viscous Transport in Eroding Porous Media

### 2019-10-10 docs/transport/erosion_transport.tex
formatting equations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Outline of the Paper}
In Section~\ref{sec:formulation}, we summarize the erosion model that is
described in more detail in previous work~\citep{qua-moo2018}.  In
Section~\ref{sec:DLP}, we recast all the governing equations as layer
potentials defined in both $\RR^2$ and in $\CC$.
Section~\ref{sec:transport} describes measures for characterizing the
geometry and transport. Section~\ref{sec:method} describes the numerical
methods, with special attention paid to the new quadrature method for
computing the shear stress and the vorticity.  Section~\ref{sec:results}
presents numerical examples for a variety of dense packings of bodies.
Finally, concluding remarks are made in Section~\ref{sec:conclusions}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Double-Layer Potential Formulation in $\RR^2$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Applying the same approach as our previous work~\citep{qua-moo2018}, we
start with the double-layer potential 
\begin{align}
  \DDD[\eeta](\xx) = \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\upi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:velocityDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  The double-layer potential is a
solution of the incompressible Stokes equations, but it cannot represent
all the solutions. Therefore, we complete the integral equation
formulation by adding the $M$ Stokeslets and rotlets~\citep{pow-mir1987}
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\upi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2}
    \right)\llambda_\ell, \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
  \quad \ell=1,\ldots,M,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$. Here, $\llambda_\ell$ and $\xi_\ell$ are the Stokeslet and
rotlet strengths, respectively, corresponding to the $\ell^{th}$ body.
Then, for any sufficiently smooth geometry $\Omega$, the solution of the
incompressible Stokes equation with a Dirichlet boundary condition $\ff$
is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:BIE}
\begin{alignat}{3}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx) +
    \NN_0[\eeta](\xx), \quad &&\xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\upi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy, &&\ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\upi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy, 
    &&\ell = 1,\ldots,M.
\end{alignat}
\end{subequations}
Here, the null space associated with the flux-free condition of $\ff$ is
addressed with  $\NN_0$ which is the integral operator with kernel
$N_0(\xx,\yy) = \nn(\xx) \otimes \nn(\yy)$, $\xx,\yy \in \Gamma$.  In
this work, $\ff$ is the prescribed velocity, which is equal to $\UU$ on
the outer wall, $\Gamma$, and equal to zero on the eroding bodies,
$\gamma_\ell$, $\ell=1,\ldots,M$.

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, we are able to
compute, for $\xx \in \Omega$, the deformation tensor, pressure, and
vorticity, defined respectively as
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\upi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  \label{eqn:pressureDLP}
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\upi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\upi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy.
\end{align}
Since the deformation tensor of the double-layer potential is
discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not exhibit any jumps across $\gamma$, so the
associated components of the deformation tensor, pressure, and vorticity
are straightforward to compute~\citep{poz1992}.  Having computed the
deformation tensor on $\gamma$, the shear stress is computed using
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In
equations~\eqref{eqn:velocityDLP},~\eqref{eqn:deformationDLP},~\eqref{eqn:pressureDLP},
and~\eqref{eqn:vorticityDLP}, the fluid velocity, deformation tensor,
pressure, and vorticity are all written as layer potentials in $\RR^2$.
However, the quadrature method we introduce in Section~\ref{sec:method}
requires a complex-valued representation of the layer potentials for the
velocity, deformation tensor, and vorticity. The first step to form a
complex representation is to write the Laplace double-layer potential as
the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\upi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Real (v(x)) \quad 
    \text{where} \quad v(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega,
  \label{eqn:laplaceComplex}
\end{align}
where $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, $\Omega$ is being interpreted as a subset of $\RR^2$ or
$\CC$. Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data of $v$
satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Having computed the boundary data of the
holomorphic function $v$, by the Cauchy integral theorem we have
\begin{align}
  \label{eqn:cauchy}
  v(x) = \frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \quad
  v'(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \quad
  v''(x) = \frac{1}{\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
\end{align}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Real (v'[\psi_1](x)) + 
      \Real (v''[y\cdot\eta](x)) - \Real (v'[\eta_1](x)) - 
      x_1\Real (v''[\eta_1](x)) - x_2\Real (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Imag (v'[\psi_1](x)) - 
      \Imag (v''[y\cdot\eta](x)) + x_1\Imag (v''[\eta_1](x)) - 
      \Real (v'[\eta_2](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Real (v'[\psi_2](x)) - 
      \Imag (v''[y\cdot\eta](x)) + \Imag (v'[\eta_1](x)) +
      x_1\Imag (v''[\eta_1](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Imag (v'[\psi_2](x)) - 
      \Real (v''[y\cdot\eta](x)) + x_1\Real (v''[\eta_1](x)) +
      \Imag (v'[\eta_2](x)) + x_2\Real (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity is
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
\Real (v'[\psi_2](x)) + \Imag (v'[\psi_1](x)) 
 + \Real (v'[\eta_2](x))+ \Imag (v'[\eta_1](x)), \quad x \in \Omega.
\end{align}

@@

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of~\citet{ioa-pap-per1991} (see equation (2.12)).  We start with
the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\upi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$,
and theses are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form simulations of nearly-touching eroding
grains, and to study dynamics of the flow in regions arbitrarily close
to eroding grains.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Outline of the Paper}
In Section~\ref{sec:formulation}, we summarize the erosion model that is
described in more detail in previous work~\citep{qua-moo2018}.  In
Section~\ref{sec:DLP}, we recast all the governing equations as layer
potentials defined in both $\RR^2$ and in $\CC$.
Section~\ref{sec:transport} describes measures for characterizing the
geometry and transport. Section~\ref{sec:method} describes the numerical
methods, with special attention paid to the new quadrature method for
computing the shear stress and the vorticity.  Section~\ref{sec:results}
presents numerical examples for a variety of dense packings of bodies.
Finally, concluding remarks are made in Section~\ref{sec:conclusions}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Double-Layer Potential Formulation in $\RR^2$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Applying the same approach as our previous work~\citep{qua-moo2018}, we
start with the double-layer potential 
\begin{align}
  \DDD[\eeta](\xx) = \int_{\bd\Omega} D(\xx,\yy) \eeta(\yy)\, ds_\yy = 
  \frac{1}{\upi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:velocityDLP}
\end{align}
where $D$ is the kernel of the integral operator, $\rr = \xx - \yy$,
$\rho = \|\rr\|$, $\nn$ is the unit outward normal at $\yy$, and $\eeta$
is an unknown density function.  The double-layer potential is a
solution of the incompressible Stokes equations, but it cannot represent
all the solutions. Therefore, we complete the integral equation
formulation by adding the $M$ Stokeslets and rotlets~\citep{pow-mir1987}
\begin{align}
  S[\llambda_\ell,\cc_\ell](\xx) = \frac{1}{4\upi} \left( 
    -\log \rho \II + \frac{\rr \otimes \rr}{\rho^2}
    \right)\llambda_\ell, \quad
  R[\xi_\ell,\cc_\ell](\xx) = \frac{\rr^\perp}{\rho^2} \xi_\ell,
  \quad \ell=1,\ldots,M,
\end{align}
where $\rr = \xx - \cc_\ell$, $\cc_\ell$ is a point inside the
$\ell^{th}$ eroding body, $\rr^\perp = (r_2,-r_1)$, and $\rho =
\|\rr\|$. Here, $\llambda_\ell$ and $\xi_\ell$ are the Stokeslet and
rotlet strengths, respectively, corresponding to the $\ell^{th}$ body.
Then, for any sufficiently smooth geometry $\Omega$, the solution of the
incompressible Stokes equation with a Dirichlet boundary condition $\ff$
is
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + 
    \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:BIE}
\begin{alignat}{3}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
    \NN_0[\eeta](\xx) \nonumber \\
    &\quad + \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx) + 
    \sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx), 
    \quad &&\qquad\xx \in \bd\Omega, \\
  \llambda_\ell &= \frac{1}{2\upi} \int_{\gamma_\ell} 
    \eeta(\yy)\, ds_\yy, &&\qquad \ell = 1,\ldots,M, \\
  \xi_\ell &= \frac{1}{2\upi} \int_{\gamma_\ell}
    (\yy - \cc_\ell)^\perp \cdot \eeta(\yy)\, ds_\yy, 
    &&\qquad \ell = 1,\ldots,M.
\end{alignat}
\end{subequations}
Here, the null space associated with the flux-free condition of $\ff$ is
addressed with  $\NN_0$ which is the integral operator with kernel
$N_0(\xx,\yy) = \nn(\xx) \otimes \nn(\yy)$, $\xx,\yy \in \Gamma$.  In
this work, $\ff$ is the prescribed velocity, which is equal to $\UU$ on
the outer wall, $\Gamma$, and equal to zero on the eroding bodies,
$\gamma_\ell$, $\ell=1,\ldots,M$.

Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, the
Stokeslets $\llambda_\ell$, and the rotlets $\xi_\ell$, we are able to
compute, for $\xx \in \Omega$, the deformation tensor, pressure, and
vorticity, defined respectively as
\begin{align}
  \label{eqn:deformationDLP} 
  \ssigma(\xx) &= \int_{\bd\Omega} \mathbf{D}_{\ssigma}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = \frac{1}{2\upi}\int_{\bd\Omega} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn) 
    \right. \nonumber \\
    & \qquad\qquad\qquad\qquad \left.
    +\frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
    \right) \, ds_\yy, \\
%
  \label{eqn:pressureDLP}
  p(\xx) &= \int_{\bd\Omega} D_{p}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\upi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) \nn \cdot
      \eeta(\yy)\,ds_\yy, \\
%
  \label{eqn:vorticityDLP} 
  \omega(\xx) &= \int_{\bd\Omega} D_{\omega}(\xx,\yy) 
      \eeta(\yy) ds_{\yy} 
  = -\frac{1}{\upi} \int_{\bd\Omega}
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4} 
      (\rr \cdot \eeta) \, ds_\yy.
\end{align}
Since the deformation tensor of the double-layer potential is
discontinuous across $\bd\Omega$, the jump term 
\begin{align}
  \frac{1}{2} \left(\pderiv{\eeta}{\ss} \cdot \ss \right) \left[
    \begin{array}{cc}
      s_x^2 - s_y^2 & 2s_x s_y \\ 2s_x s_y & s_y^2 - s_x^2
    \end{array}
  \right],
  \label{eqn:deformationJump}
\end{align}
must be added to $\ssigma(\xx)$ for $\xx \in \bd\Omega$.  Finally, the
Stokeslets and rotlets do not exhibit any jumps across $\gamma$, so the
associated components of the deformation tensor, pressure, and vorticity
are straightforward to compute~\citep{poz1992}.  Having computed the
deformation tensor on $\gamma$, the shear stress is computed using
equation~\eqref{eqn:shearStress}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Double-Layer
Potential}
\label{sec:DLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In
equations~\eqref{eqn:velocityDLP},~\eqref{eqn:deformationDLP},~\eqref{eqn:pressureDLP},
and~\eqref{eqn:vorticityDLP}, the fluid velocity, deformation tensor,
pressure, and vorticity are all written as layer potentials in $\RR^2$.
However, the quadrature method we introduce in Section~\ref{sec:method}
requires a complex-valued representation of the layer potentials for the
velocity, deformation tensor, and vorticity. The first step to form a
complex representation is to write the Laplace double-layer potential as
the complex integral
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{2\upi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2}\eeta(\yy)\, ds_\yy = \Real (v(x)),
\end{align}
where
\begin{align}
  v(x) = \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy, \quad x \in \Omega.
  \label{eqn:laplaceComplex}
\end{align}
Here $x = x_1 + i x_2,y = y_1 + i y_2 \in \CC$ are the complex
counterparts of $\xx = (x_1,x_2),\yy = (y_1,y_2) \in \RR^2$, and $\eta =
\eta_1 + i \eta_2$ is the complex counterpart of $\eeta =
(\eta_1,\eta_2)$. Therefore, depending on the formulation of the layer
potential, $\Omega$ is being interpreted as a subset of $\RR^2$ or
$\CC$. Equation~\eqref{eqn:laplaceComplex} is converted to a Cauchy
integral by first finding the boundary data of $v$.  Assuming that
$\Omega$ is a simply-connected interior domain, the boundary data of $v$
satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  \label{eqn:SPrelation}
  v(x) = - \frac{1}{2} \eta(x) + \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x-y}\, dy, \quad x \in \Omega.
\end{align}
For exterior domains, the jump term changes from $-1/2$ to $1/2$, and
for multiply-connected domains, such as a porous media, $\bd\Omega$ is
decomposed into its different connected components and the appropriate
jump relation is applied.  Having computed the boundary data of the
holomorphic function $v$, by the Cauchy integral theorem we have
\begin{subequations}
  \begin{alignat}{3}
  \label{eqn:cauchy}
  v(x) &= \frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{v(y)}{y-x} \,dy, \\
  v'(x) &= \frac{1}{2\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^2} \, dy, \\
  v''(x) &= \frac{1}{\upi i} \int_{\bd\Omega}
    \frac{v(y)}{(y-x)^3} \, dy,
  \end{alignat}
\end{subequations}
for $x \in \Omega$.  Since $v(x)$ depends on the complex-valued density
function $\eta$, we use the notation $v[\eta](x)$ for the holomorphic
function defined in equation~\eqref{eqn:laplaceComplex}, and its first
two derivatives are written as $v'[\eta](x)$ and $v''[\eta](x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation for the Gradient of the
Double-Layer Potential}
\label{sec:gradDLPcomplex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Computing the shear stress and vorticity requires a complex-valued layer
potential representation of the velocity gradient.  The deformation
tensor at $x \in \Omega$ is found by computing the derivatives of the
expressions for $u_1$ and $u_2$ in equation~\eqref{eqn:cauchyVelocity}  
\begin{equation}
\label{eqn:cauchyGradient}
  \begin{aligned}
    \pderiv{u_1}{x_1} &= +\Real (v'[\psi_1](x)) + 
    \Real (v''[y\cdot\eta](x)) - \Real (v'[\eta_1](x)) \\
    &- x_1\Real (v''[\eta_1](x)) - x_2\Real (v''[\eta_2](x)), \\
    \pderiv{u_1}{x_2} &= - \Imag (v'[\psi_1](x)) - 
    \Imag (v''[y\cdot\eta](x)) + x_1\Imag (v''[\eta_1](x)) \\
    &- \Real (v'[\eta_2](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_1} &= +\Real (v'[\psi_2](x)) - 
    \Imag (v''[y\cdot\eta](x)) + \Imag (v'[\eta_1](x))  \\
    &+ x_1\Imag (v''[\eta_1](x)) + x_2\Imag (v''[\eta_2](x)), \\
    \pderiv{u_2}{x_2} &= -\Imag (v'[\psi_2](x)) - 
    \Real (v''[y\cdot\eta](x)) + x_1\Real (v''[\eta_1](x)) \\
    &+ \Imag (v'[\eta_2](x)) + x_2\Real (v''[\eta_2](x)).
  \end{aligned}
\end{equation}
The same expressions are used to compute the deformation tensor for $x
\in \bd\Omega$, except that the jump
condition~\eqref{eqn:deformationJump} is included.  Finally, to compute
the shear stress, the deformation tensor on $\bd\Omega$ is applied to
the normal and tangent vectors as in equation~\eqref{eqn:shearStress}.
The velocity gradient is also used to compute the vorticity in the fluid
bulk.  For $x \in \Omega$, the Cauchy integral representation of the
vorticity at $x \in \Omega$ is
\begin{equation}
  \begin{aligned}
    \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} &= 
      \Real (v'[\psi_2](x)) + \Imag (v'[\psi_1](x))  \\
      &+ \Real (v'[\eta_2](x))+ \Imag (v'[\eta_1](x)).
  \end{aligned}
\end{equation}

@@

To compute the shear stress and vorticity, we require a Barycentric
quadrature rule for $v''(x)$.  The derivation is largely based on the
work of~\citet[see equation (2.12)]{ioa-pap-per1991}.  We start with
the second derivative of the Cauchy integral theorem
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x).
\end{align}
For the interior case, $x \in \Omega$, we use the identity
\begin{align}
  \frac{1}{2\upi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n = 2,3,\ldots. \\
    \end{array}
  \right.
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
This integrand is constructed so that it is bounded for all $x \in
\Omega$, and applying the trapezoid rule, we have
\begin{align}
  0 \approx  \sum_{j=1}^{N} 
      \frac{2v^{-}(y_j) - v''(x)(y_j-x)^2 - 2v(x) - 2(y_j-x)v'(x)}
      {(y_j-x)^3} w_j,
\end{align}
where the accuracy is independent of $x$.  Solving for $v''(x)$, the
Barycentric quadrature rule for the interior second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}, \quad x \in \Omega.
\end{align}
For the exterior case, $x \in \Omega^c$, we start with the identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\upi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x}\, dy. 
\end{align}
Combining this identity with the Cauchy integral representation of
$v''(x)$, we have
\begin{align}
  0 = \frac{1}{2\upi i} \int_{\bd\Omega} 
    \frac{2v^+(y) - 2v(x) - 2(y-x)v'(x) + (y-a)^{-1} (x-a) (y-x)^2 v''(x)}
    {(y-x)^3}.
\end{align}
As in the interior case, the integrand is chosen so that it is bounded
for all $x \in \Omega^c$.  Therefore, after applying the trapezoid rule
and solving for $v''(x)$, we have the Barycentric quadrature rule for
the exterior second derivative
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}, 
    \quad x \in \Omega^c.
\end{align}
The quadrature rule for $v''(x)$ requires the values $v(x)$ and $v'(x)$,
and theses are computed using the quadrature rules in
equations~\eqref{eqn:BaryvInterior},~\eqref{eqn:BaryvExterior},~\eqref{eqn:BaryvprimeInterior},
and~\eqref{eqn:BaryvprimeExterior}. In Section~\ref{sec:results}, these
quadrature rules are used to form simulations of nearly-touching eroding
grains, and to study dynamics of the flow in regions arbitrarily close
to eroding grains.
